### 2017-05-31 sheets/talk-4/sheet4.tex
fixed ex.1a

\begin{document}
\[\uebunghead{Talk 4}{Galois Connections and Adjunctions}
\author {Sarah Mameche}
\section{Galois connections}

%Solutions preliminary

\begin{exercise}
Let $\mathscr{P} = (P, \preceq_P)$ and $\mathscr{Q} = (Q, \preceq_Q)$ be partially ordered sets, and $F \from P \to Q, G \from Q \to P$ monotone functions with $F \dashv G$.
\begin{enumerate}
\item Show that the following characterizations of the Galois connection are equivalent.
\begin{enumerate}
\item $p \preceq_P G (F (p)) \land q \preceq_Q F(G (q))$
\item $p \preceq_P G(q) \Leftrightarrow q \preceq_Q F(p)$
\end{enumerate}
\item Show that the following property holds for $F \dashv G$. 
\[  
\begin{aligned}
 \forall q \in Q.\; G(q) = min\{p \in P \such F(p) \preceq_Q q\}
\end{aligned}
\]
Formulate a similar statement $ \forall p \in P, F(p)$.
\end{enumerate}
\end{exercise}

\begin{answer}
\begin{enumerate}
\item ...
\item %For some $q \in Q$, set $S = \{p \in P \such F(p) \preceq_Q q}.
Dual property: $\forall p \in P.\; F(p) = min\{q \in Q \such p \preceq_P G(q)\} $
\end{enumerate}
\end{answer}

@@

\begin{answer}
\begin{enumerate} \item right adjunct: floor function  $\mathcal{F} : \mathcal{Q} \to  \mathcal{N}$ \\
left adjunct: ceiling function $\mathcal{C} : \mathcal{Q} \to  \mathcal{N}$
\item There can be no galois connection $F \dashv G$ from $\mathscr{Q}$ to $\mathscr{N} . F(q)=1 \Leftrightarrow 1 q<2,$ so $\{q\in \mathbb{Q}^+ \such F(q) = 1\} $ has no maximum. The statement follows with Ex.1b
\end{enumerate}\end{answer}

\section{Adjunctions}
\begin{exercise}
Recall the naturality axiom that has to hold for the isomorphism $\mathscr{B}(F(A), B) \; \cong \; \mathscr{A}(A, G(B)) $:

@@

\begin{answer}
(i), (ii) $\Rightarrow$ (iii)\\
composing any $q \from B \to B'$ with (ii): $\forall f, p: \forall q: q \of \bar{f} \of F(p) = q \of \overline{f \of p} $ (by ii). \\Since $f \from p \to F(A') \to B$, its transpose $h=:\overline{f \of p} \from A' \to B.$ \\So from (i) and quantif. over A: $q \of \overline{f \of p}  = q \of h = \overline{G(q) \of \bar{h}} = \overline{G(q) \of f \of p}$\\
(iii) $\Rightarrow$ (i)\\
%(iii) quantifies over all $p$, so set $p = id_A$ (exists for each object). Then $f \of p \from A \to G(B)$, and $\bar{f} \of F(p) \from F(A) \to B.$ 
Assume (i) is false. 
Taking a $g \from F(A) \to B$, you get $h = g  \of id_B  \from F(A) \to B$. 
Then the equality (i) does not hold for h, that is \\ 
$\forall q, \overline{q \of h} \not= G(q) \of \bar{h}$ 
\\$ \Leftrightarrow \overline{q \of g  \of id_B} \not= G(q)  \of \overline{h}$ ( Def h)\\
$\Leftrightarrow q \of g  \of id_B \not= \overline{G(q)  \of \overline{h}}$ (Def transposition)

$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of \overline{h}} $  (define f as $\bar {g}.$ Also, $  F(id_A) = id_B$ by Def of functors)
\\$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of \overline{g}} \; (h = g \of id_B = g)$
\\$\Leftrightarrow q \of \bar{f}  \of F(id_A) \not= \overline{G(q)  \of f \of id_A}\;  (f = \bar{g} $ by def. and $f = f \of id_A)$
\\So (iii) is false for $f = \bar {g}$ and $ p= id_A$. \\(iii)  $\Rightarrow$ (i) similarly.
\end{answer}

\begin{exercise}
Find a right adjunct to the forgetful functor $F \from $\textbf{Set }$ \to $\textbf{ Rel}.
\end{exercise}
\begin{answer}
\end{answer}

@@

\begin{answer}
\begin{enumerate}
\item We have seen that initial (terminal) objects of categories correspond to left (right) adjoints. So for any category that has a zero object, there is a functor that is both left and right adjoint. Also, an equivalence between two categories is both a left and right adjoint.
\item The contravariant powerset functor $\mathcal{P} \from $\textbf {Set }$ \to$\textbf{ Set}$^{op}$ (see Ex. 7) is self-adjoint. Functions $A \to \mathcal{P}(B)$ in \textbf{Set} correspond to subsets of $A \times B$ and thus to functions $B \to \mathcal{P}(A)$.
\end{enumerate}
\end{answer}

@@

\begin{answer}
Assume $I$ is initial in $ \mathscr{A}$, then maps $I \to A$ are unique $\forall A \in A$. This implies that maps  $I \to G(B)$ are unique $\forall B$ because $G(B) \in ob(\mathscr{A})$. From $F\dashv G$, we get that the corresponding maps $F(I) \to B$ are unique $\forall B$, so $F(I)$ initial in $\mathscr{B}$.\\ It can be shown in the same way that terminal objects are preserved by $G$.\end{answer}

@@

\begin{answer}
$F \dashv F^{-1}$, meaning $F(f)(X) \subseteq Y \Leftrightarrow X \subseteq F^{-1}$.\\
Assume $F \dashv F^{-1}$. Then $\forall x \in X, f(x) \in Y$ and thus $X \subseteq F^{-1}$\\
Assume $X \subseteq F^{-1}$ and take $b \in F^{-1}$. By Def $f(x) = b$ for some x, and with $X \subseteq F^{-1}, b = f(x) \in Y$\\
Groups: f homomorphism, $\mathcal{P}(G)$ set of subgroups, same proof \\
Top. Sp: continuous maps and subspaces

\end{answer}

@@

\begin{answer}
There is a left adjoint $D$, where $D(A)$ is a discrete category with objects $A$. There is also a right adjoint $I$, where $I(A)$ is the category with objects $A$ and a unique arrow $a \to b$ for each $a,b \in A$.
\end{answer}

@@

\]
\end{document}

>>>>

\begin{document}
	\uebunghead{Talk 4}{Galois Connections and Adjunctions}
	\author {Sarah Mameche}

	\section{Galois connections}

	\begin{exercise}
		Let $\mathscr{P} = (P, \preceq_P)$ and $\mathscr{Q} = (Q, \preceq_Q)$ be partially ordered sets, and $F \from P \to Q, G \from Q \to P$ monotone functions with $F \dashv G$.
		\begin{enumerate}
			\item Show that the following characterizations of the Galois connection are equivalent.
			\begin{enumerate}
				\item $p \preceq_P G (F (p)) \land F(G (q)) \preceq_Q q$
				\item $p \preceq_P G(q) \Leftrightarrow F(p) \preceq_Q q$
			\end{enumerate}
			\item Show that the following property holds for $F \dashv G$. 
			\[  
			\begin{aligned}
			\forall q \in Q.\; G(q) = max\{p \in P \such F(p) \preceq_Q q\}
			\end{aligned}
			\]
			Formulate a similar statement $ \forall p \in P, F(p)$.
		\end{enumerate}
	\end{exercise}

	\begin{answer}
		\begin{enumerate}
			\item (i) $\Rightarrow$ (ii):
			\\Assume (i) and $p \preceq_P G(q)$. Since $F$ is monotone, $F(p) \preceq_Q F(G(q))$, and with (i):  $F(p) \preceq_Q F(G(q)) \preceq_Q q$.
			\\Assume (i) and $F(p) \preceq_Q q$. Applying monotone $G$ and using (i), $ p \preceq_P G(F(p)) \preceq_P G(q)$.\\ \\
			(ii) $\Rightarrow$ (i): \\Let $q := F(p)$. Since $F(p) \preceq_Q F(p)$, (ii) implies $p \preceq_P G (F (p))$. 
			\\Similarly, with $p := G(q)$,  (ii) implies $F(G (q)) \preceq_Q q$.

			\item  For some $q \in Q$, set $D = \{p \in P \such F(p) \preceq_Q q\}$. 
			\\We know from (a) that $\forall q, F(G(q)) \preceq_Q q$, so $G(q) \in D$.
			\\To show that $G(q)$ is also the maximum of D, take some $d \in D$, i.e. $F(d) \preceq_Q q$. 
			\\With monotonicity of G, $G(F(d)) \preceq_Q G(q)$. From (a), $d \preceq_Q G(F(d))$, so $d \preceq_Q G(q)$.
			\\
			\\Dual property: $\forall p \in P.\; F(p) = min\{q \in Q \such p \preceq_P G(q)\} $
		\end{enumerate}
	\end{answer}

@@

	\begin{answer}
		\begin{enumerate} 
			\item The right adjunct $G \from \mathbb{Q}^+ \to  \mathbb{N}$ is the floor function, because for $x \in \mathbb{N}, y \in \mathbb{Q}^+:$
			\\$ x = \lfloor x \rfloor = G(I(x))  \land I(G(y)) =  \lfloor y \rfloor \leqslant y$.
			\\The left adjunct $H \from \mathbb{Q}^+ \to  \mathbb{N}$ is the ceiling function.\\ For $x \in \mathbb{N}, y \in \mathbb{Q}^+: y \leqslant \lceil y \rceil = I(H(x))  \land H(I(x)) =  \lceil x \rceil = x$.\\

			\item $I \dashv G$ but not $G \dashv I$. G has no right adjoint:
			\\$\{q\in \mathbb{Q}^+ \such G(q) \leqslant 1\} $ has no maximum, because $G(q)= 1 \; \forall 1 \leqslant q < 2$. 
			The statement follows with Ex.1 (b).
		\end{enumerate}
	\end{answer}

	\section{Adjunctions}

	\begin{exercise}
		Recall the naturality axiom that has to hold for the isomorphism $\mathscr{B}(F(A), B) \; \cong \; \mathscr{A}(A, G(B)) $:

@@

	\begin{answer}
		(i), (ii) $\Rightarrow$ (iii)
		\\$\forall q \from B \to B', \forall f, p. \;$\\$ q \of \bar{f} \of F(p) = q \of \overline{f \of p}\;\;$ ((ii) and naturality of composition)\\
		$\Rightarrow q \of \bar{f} \of F(p)  = q \of h\;\;$ (*)\\
		$\Leftrightarrow  q \of \bar{f} \of F(p) = \overline{G(q) \of \bar{h}}\;\;$ (i)\\
		$\Leftrightarrow  q \of \bar{f} \of F(p)= \overline{G(q) \of f \of p}\;\;$(*)\\
		In (*), $F \dashv G$ is used, so the function  $\overline{f \of p} \from A' \to G(B)$ has a transpose $ h =: f \of p \from F(A') \to B$.\\\\
		(iii) $\Rightarrow$ (i)\\
		Assume (iii) and let $p = id_A. \; F(p) = id_B$ by the functor axioms. Due to $F \dashv G$, every $f\from A \to G(B)$ has a transpose $g := \bar{f}$. Then \\\\
		$\forall q, f. \; \overline{G(q) \of f \of p} = q \of \bar{f} \of F(p)$\\
		$\Rightarrow \forall q, g.  \;\overline{G(q) \of \bar{g} \of id_A} = q \of g \of id_B$ (instantiations) \\
		$\Leftrightarrow \forall q, g.  \;\overline{G(q) \of \bar{g}} = q \of g$ (identities)\\
		$\Leftrightarrow \forall q, g. \; G(q) \of \bar{g} = \overline{q \of g}$ (transposition)\\

		(iii)  $\Rightarrow$ (ii) analogously.
	\end{answer}

	\begin{exercise}
		Find a right adjunct to the forgetful functor $F \from $\textbf{Set }$ \to $\textbf{ Rel}.
	\end{exercise}

	\begin{answer}
		The right adjunct is the powerset functor $P \from $\textbf{Rel }$ \to $ \textbf{ Set}. 
		It maps a set $A$ to its powerset $\mathcal{P}(A)$. 
		A Relation $R \subseteq A \times B$ is mapped to the function $f_R \from \mathcal{P}(A) \to \mathcal{P}(B), f_R(X) = \{b \in B \such \exists a \in X. (a,b) \in R \}$. 
		\\Every function $f \from A \to \mathcal{P}(A)$ can be seen as a relation $R_f \subseteq F(A) \times B$, with $(x,y) \in R \Leftrightarrow y \in f(x)$, and vice versa. 
		So \textbf{Rel }$(F(A), B) \cong $\textbf{ Set}$(A, P(B))$.\\
		Naturality in $A$: for functions $h \from A' \to A$,  the constructed relation for $f \of h$ is $R_{f \of h} = \{(x,y) \such  y \in f(h(x))\}.$ 
		The relation $R_f \of P(h)$ is the same one:  $R_f \of P(h)= R_f \of R_h = \{(x,z) \such \exists y. \; z \in f(y) \land y \in h(x)\} = R_{f \of h}$. 
		\\Naturality in $B$: for relations $R \subseteq B \times B', f_{B \times B'}$ is the same as $f_B \of F(B').$
	\end{answer}

@@

	\begin{answer}
		\begin{enumerate}
			\item We have seen that initial (terminal) objects of categories correspond to left (right) adjoints. So for any category that has a zero object, there is a functor that is both left and right adjoint. Also, an equivalence between two categories is both a left and right adjoint.

			\item The contravariant powerset functor $\mathcal{P} \from $\textbf {Set }$ \to$\textbf{ Set}$^{op}$ (see Ex. 7) is self-adjoint. 
			As seen in Exercise 4, Functions $A \to \mathcal{P}(B)$ in \textbf{Set} correspond to subsets of $A \times B$ and thus also to functions $B \to \mathcal{P}(A)$.
		\end{enumerate}
	\end{answer}

@@

	\begin{answer}
		Assume $I$ is initial in $ \mathscr{A}$, then $\forall A \in ob(\mathscr{A})$, the maps $I \to A$ are unique. 
		This implies that $\forall B \in ob(\mathscr{B})$, the maps  $I \to G(B)$ are unique because $G(B) \in ob(\mathscr{A})$. 
		From $F\dashv G$, we get that the corresponding maps $F(I) \to B$ are unique, so $F(I)$ initial in $\mathscr{B}$.\\\\
		Dually, terminal objects are also preserved by $G$.
	\end{answer}

@@

	\begin{answer}
		$F \dashv F^{-1}$, meaning $F(X) \subseteq Y \Leftrightarrow X \subseteq F^{-1}(Y)$.\\
		Assume $F(X) \subseteq Y$.  $\forall x \in X, f(x) \in Y$ and thus $X \subseteq F^{-1}(Y)$\\
		Assume $X \subseteq F^{-1}(Y)$, then $\forall x \; x \in F(Y)$, which means $F(X) \subseteq Y$\\
		For Groups, f is a homomorphism, and $\mathcal{P}(G)$ the set of subgroups.\\
		For Topological Spaces, we have continuous maps and subspaces.
	\end{answer}

@@

	\begin{answer}
		There is a left adjoint $D$, where $D(A)$ is a discrete category with objects $A$ (for some set $A$). \\
		For a category $\mathscr{B}$, the functors in \textbf{CAT} $(D(A), \mathscr{B})$ are fully defined by the object mapping, since $D(A)$ has only identity arrows. So the functors can be seen as functions in \textbf{Set} $(A$, ob($\mathscr{B}))$.
		\\There is also a right adjoint $I$, where $I(A)$ is the category with objects $A$ and a unique arrow $a \to b$ for each $a,b \in A$. Here, functors in \textbf{CAT} $(\mathscr{B}, I(A))$ can also be seen as functions in \textbf{Set} (ob($\mathscr{B}), A)$. That is because a functor $F$ must map arrows $B \to B'$ to $F(B) \to F(B')$, and if there is a unique arrow $F(B) \to F(B')$, we need only the object mapping to describe the functor.
	\end{answer}

@@

\end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
